Minimax optimal Laplacian smoothing

### 2021-02-11 notes/manuscripts/laplacian_smoothing/AISTATS/archive/minimax_optimal_laplacian_smoothing.tex
Tampered with directory structure.

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}
\newcommand{\lambdavec}{\boldsymbol{\lambda}}

@@

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\SM}{\mathrm{SM}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}

@@

\begin{document}
\title{Minimax optimal Laplacian smoothing}
\author{Alden Green, Sivaraman Balakrishnan, and Ryan J. Tibshirani}
\date{\today}
\maketitle

\begin{abstract}
	\textcolor{red}{(TODO)}:
\end{abstract}

@@

In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Reals^d$, and 
\begin{equation}
\label{eqn:random_design_regression}
Y_i = f_0(X_i) + \varepsilon_i
\end{equation}
with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise. We will assume the following throughout:
\begin{enumerate}[label=(P\arabic*)]
	\item
	\label{asmp:bounded_lipschitz_density} 
	$P$ admits a density $p$ bounded away from $0$ and $\infty$, i.e.
	\begin{equation*}
	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty,~~\textrm{for all $x \in \Xset$.}
	\end{equation*}
	Additionally, $p$ is Lipschitz, with Lipschitz constant $L_p$.
\end{enumerate}
Our goal is to perform statistical inference on the unknown regression function $f_0: \Xset \to \Reals$, by which we mean either (a) \emph{estimating} $f_0$ by $\wh{f}$, an estimator constructed from the data $(X_1,Y_1),\ldots,(X_n,Y_n)$ or (b) simply \emph{testing} whether $f_0 = 0$, i.e whether there is any signal present.

In graph-based nonparametric regression, we perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense, and then constructing an estimate $\wh{f}$ which is smooth with respect to the graph $G_{n,r}$. The neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$. The $n \times n$ weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ encodes proximity between pairs of samples; for a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the entries $\mathbf{W}_{ij}$ are given by
\begin{equation*}
\label{eqn:neighborhood_graph}
{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}}{r}\Biggr).
\end{equation*}
where $\norm{\cdot} = \norm{\cdot}_{\Rd}$ is the Euclidean norm. Then the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$, and the graph Laplacian can be written as
\begin{equation}
\label{eqn:graph_Laplacian}
\Lap_{n,r} = \bf{D} - \bf{W}
\end{equation}

The Laplacian smoothing estimator $\wh{f}_{\LS}$ \citep{smola2003} is a penalized least squares estimator, given by
\begin{equation}
\label{eqn:laplacian_smoothing}
\wh{f}_{\LS} := \min_{f \in \Reals^n} \biggl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_{n,r} f \biggr\}
\end{equation}
where $\rho > 0$ acts a tuning parameter on the penalty $f^T \Lap_{{n,r}} f$.
Assuming~\eqref{eqn:laplacian_smoothing} is a reasonable estimator of $f_0$, the squared empirical norm
\begin{equation}
T_{\LS} = \frac{1}{n}\bigl\|\wt{f}_{\LS}\bigr\|_2^2 \label{eqn:laplacian_smoothing_test}
\end{equation}
is in turn a reasonable statistic to assess whether or not $f_0 = 0$.

Laplacian smoothing has many advantages compared to other nonparametric regression methods. For instance:
\begin{itemize}
	\item It is fast, easy, and stable to compute. Letting ${\bf Y} = (Y_1,\ldots,Y_n) \in \Reals^n$, it is not hard to see that the minimizer $\wt{f}_{\LS} = (\rho \Lap_{n,r} + I)^{-1}{\bf Y}$ can be computed by solving a linear equation; in practice, one typically chooses $r$ to be small, with the result being that the matrix $\Lap_{n,r}$ is sparse. There exist known fast solvers of exactly this system.
	\item Is is generalizable to non-standard data---for instance, text or images, or really any data modality on which it is possible to define a kernel.
	\item It is easily adaptable to the semi-supervised learning setting. 
\end{itemize}
\textcolor{red}{(TODO)}: Perhaps additional motivation is needed?

\paragraph{Related work.} For this reason a considerable body of work has emerged analyzing the properties of Laplacian smoothing, as well as of graph-based methods more generally. Roughly speaking this work can divided into two categories.

Works in the first category view the responses $Y = \theta + \varepsilon$ as noisy samples of a regression function $\theta \in \Reals^n$ defined on the vertices of a \emph{fixed graph} $G$. and bound the statistical error as a function of $\theta$ and $G$. (In the estimation literature, see for instance \citep{wang2016, sadhanala16,sadhanala17,kirichenko2017,kirichenko2018}; in the testing literature, see \citep{sharpnack2013,sharpnack2013b,sharpnack2015}.) Note that in our setting, the graph $G_{n,r}$ and the evaluations $(f_0(X_1),\ldots,f_0(X_n))$ are random rather than fixed, and our assumptions are placed only on the underlying objects $P$ and $f_0$ from which these random object arise. As such, both our analysis and conclusions are naturally quite different.

Works in the second category prove the consistency of neighborhood graph Laplacian smoothing and other neighborhood graph-based methods in the large-sample limit. They do so by establishing the convergence of the graph Laplacian $\Lap_{{n,r}}$ to the continuum Laplace operator $\Delta_P: C^2(\Xset) \to \Leb^2(\Xset)$,
\begin{equation}
\label{eqn:laplace_operator}
\Delta_Pg := -\frac{1}{p} \dive(p^2 \nabla g)
\end{equation}
as $n \to \infty$ and $r \to 0$. This is meant in (at least) three senses: (1) \emph{pointwise convergence}, meaning that $\Lap_{n,r}f \to \Delta_Pf$ in an appropriate norm \citep{belkin03,belkin05,lafon04,hein05,singer06,gine06}; (2) spectral convergence, meaning the eigenvalues $\lambda_k(\Lap_{n,r}) \to \lambda_k(\Delta_P)$, where the latter are the Dirichlet eigenvalues of $\Delta_P$ \citep{belkin07,burago2014,trillos2018,trillos2019,calder2019}.\footnote{\label{footnote:dirichlet_eigenvalues} Formally, Dirichlet eigenvalues are solutions to the problem
	$$
	\Delta_Pg = \lambda g,\\
	\restr{g}{\partial \Xset} = 0
	$$}, and (3) convergence of norms, meaning $f^T \Lap_{n,r} f \to \langle f,\Delta_Pf \rangle_{\Leb^2(P)}$ \citep{bousquet03,hein06,zhou11}. While these results imply \emph{consistency} of many different graph Laplacian based procedures, the statistical \emph{optimality} of Laplacian smoothing over neighborhood graphs is still not well understood.

\paragraph{Our contributions.} Our three main contributions lie in this latter direction, and they can all be summarized as follows: Laplacian smoothing methods are minimax optimal over first order Sobolev spaces. In more detail, we will show that when $f_0$ belongs to the first order Sobolev ball $H^1(\Xset;M)$:
\begin{itemize}
	\item \textbf{Minimax optimal estimation.} With high probability, $\frac{1}{n}\bigl\|\wt{f}_{\LS} - f_0\bigr\|_{2}^2 \lesssim M^{d/(2+d)}n^{-2/(2 + d)}$.
	\item \textbf{Minimax optimal testing.}
	A level-$\alpha$ test constructed using $T_{\LS}$ has non-trivial power whenever $\norm{f_0}_{\Leb^2(\Xset)}^2 \gtrsim M^{2d/(4 + d)}n^{-4/(4 + d)}$. 
	\item \textbf{Manifold adaptivity.}
	If $\mc{X} \subset \Rd$ is a submanifold of dimension $m < d$, both of the aforementioned rates hold with $d$ replaced by $m$.
\end{itemize}

\paragraph{Organization.} We now describe the structure of the rest of this paper, which builds to the aforementioned minimax results.
\begin{itemize}
	\item In Section~\ref{sec:minimax_optimal_regression_sobolev_spaces}, for completeness we review some standard results regarding minimax nonparametric regression over Sobolev spaces.
	\item In Section~\ref{sec:graph_sobolev_classes}, we introduce the graph Sobolev class $H^1(G_{n,r})$, which we view as a graph-based analog to the continuum Sobolev class $H^1(\Xset)$. We show that key properties of $H^1(\Xset)$ used to derive minimax rates are shared by $H^1(G_{n,r})$.
	\item In Section~\ref{sec:minimax_optimal_laplacian_smoothing}, we leverage the results of Section~\ref{sec:graph_sobolev_classes} to derive tight upper bounds on the error of Laplacian smoothing methods.
	\item In Section~\ref{sec:simulations} we provide some experiments supporting our theory, before concluding in Section~\ref{sec:discussion}.
\end{itemize}

\textbf{Notation.}

We use $A,A_1, \ldots$ and $C,C_1, \ldots$ to refer to constants which do not depend on $n$, $r$, or $f_0$; likewise with $a, a_1,\ldots$ and $c, c_1,\ldots$.

\textcolor{red}{(TODO)}: Complete this.

\section{Minimax-optimal regression over Sobolev spaces}
\label{sec:minimax_optimal_regression_sobolev_spaces}

Before we get to our main results, we review minimax optimal estimation and testing rates over Sobolev classes. We also single out a notable method which achieves these rates: \emph{smoothing splines}. As we will see, Laplacian smoothing can be seen as a graph-based---i.e. discrete---counterpart to the ``continuous-time'' approach of smoothing splines.

\paragraph{Sobolev spaces, and the Sobolev embedding theorem.} We start by briefly introducing Sobolev spaces, emphasizing only the material relevant to our statistical context.

Roughly speaking, (Hilbert-)Sobolev spaces contain functions $f \in \Leb^2(\Xset)$ with derivatives which themselves belong to $\Leb^2(\Xset)$. More formally, for a given integer $s$, the Sobolev space $H^{s}(\Xset)$ consists of all functions $f \in \Leb^2(\Xset)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d) \in \mathbb{N}^d$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ exists and belongs to $\Leb^2(\Xset)$. For such functions, the Sobolev seminorm $\seminorm{f}_{H^{s}(\Xset)}$ and norm $\norm{f}_{H^{s}(\Xset)}$ are given by 
\begin{equation*}
\seminorm{f}_{H^s(\Xset)}^2 = \sum_{\abs{\alpha} = s}\int_{\mathcal{X}} \Bigl|\bigl(D^{\alpha}f\bigr)(x)\Bigr|^2 \,dx, ~~ \norm{f}_{H^{s}(\Xset)}^2 = \sum_{k = 0}^{s} \seminorm{f}_{H^k(\Xset)}^2
\end{equation*}
and for a given $M > 0$, the order $s$ Sobolev ball is $H^s(\Xset, M) = \set{f: \norm{f}_{H^s(\Xset)} \leq M}$. The graph Sobolev seminorm $f^T \Lap_{n,r} f$ should be viewed as a random, discrete approximation to the continuum first-order Sobolev seminorm $|f|_{H^1(\Xset)}^2$, and so in our analysis of Laplacian smoothing we will mainly focus on the first-order Sobolev space $H^1(\Xset)$ hereafter.

Much of statistical analysis of nonparametric regression over Sobolev spaces relies on the Sobolev embedding theorem, which establishes that functions in Sobolev spaces also satisfy other, (even more) classical notions of smoothness. Concretely, the Sobolev space $H^1(\Xset)$ will continuously embed either into a Holder space $C^{0,q}(\Xset)$ or an $\Leb^p$ space, depending on the dimension $d$.
\begin{itemize}
	\item When $d = 1$, any function $f \in H^1(\Xset)$ satisfies $\|f\|_{H^1(\Xset)} \leq C\|f\|_{C^{0,q}(\Xset)}$ for $q = 1/2.$
	\item When $d > 2$, any function $f \in H^1(\Xset)$ satisfies $\|f\|_{H^1(\Xset)} \leq C\|f\|_{\Leb^{p}(\Xset)}$ for $p = 2d/(d - 2)$. 
\end{itemize}
In the above, the exponents $p$ and $q$ cannot be improved upon.

% AJG commented this out.
% When $d = 1$, the embedding of $H^1(\Xset)$ into $C^{0,q}(\Xset)$ facilitates a general analysis of (penalized) least estimators over $H^1(\Xset)$, using arguments from empirical process theory. However when $d \geq 2$, such arguments fall apart: for example, good bounds on the metric entropy of $H^1(\Xset)$ are not known in this setting. More troubling still, least squares estimators over $H^1(\Xset)$ are not even well-posed when $d \geq 2$.

% For these reasons, analysis of least squares estimators for the nonparametric regression problem over Sobolev spaces typically assume $d = 1$---or more generally that $f_0 \in H^s(\Xset)$ for some $2s > d$. However, the Laplacian smoother $\wh{f}_{\LS}$ is a special least-squares estimator---it has a nice closed-form solution amenable to a bias-variance decomposition. We therefore do not need to rely on empirical process theory, and so we will be able to analyze $\wh{f}_{\LS}$ and $T_{\LS}$ beyond the univariate setting.

As we will see, the Sobolev embedding theorem has rather grave consequences for least-squares estimators over $H^1(\Xset)$ when $d > 1$, rendering such estimators ill-posed. However, the Laplacian smoother $\wh{f}_{\LS}$ is a special least-squares estimator. It has a nice closed-form solution amenable to a bias-variance decomposition, and so we will be able to analyze $\wh{f}_{\LS}$ and $T_{\LS}$ beyond the univariate setting.

\paragraph{Connecting Laplacian smoothing to smoothing splines.} 
To better motivate Laplacian smoothing as a tool for nonparametric regression, we connect it to one of the most classical methods of nonparametric regression: smoothing splines.

Smoothing splines are specially designed for regression over Sobolev spaces, as they are penalized least squares estimators where the penalty is itself a Sobolev semi-norm. Mathematically, the (first-order) smoothing spline estimator $\wh{f}_{\SM}$ is defined as the solution to the following optimization problem,
\begin{equation}
\label{eqn:smoothing_spline}
\wt{f}_{\SM} = \argmin_{f} \biggl\{\sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \rho \cdot \seminorm{f}_{H^{1}(\Xset)}^2\biggr\}.
\end{equation}
In~\eqref{eqn:smoothing_spline}, the minimum can either be taken over all $f \in H^1(\Xset)$ or over all $f \in C^1(\Xset)$; we shall refer to the former as smoothing splines, and the latter as \emph{Lipschitz} smoothing splines.

By viewing $\Delta_P$ as the continuum limit of~$\Lap_{n,r}$, one can tie the penalty used by the smoothing spline estimator $\wt{f}_{\SM}$ to that used by the Laplacian smoothing estimator $\wh{f}_{\LS}$. Suppose for the moment that $P$ is the uniform distribution over $\Xset$. Using the definition~\eqref{eqn:laplace_operator}, integrating by parts, and invoking the aforementioned convergence properties of $\Lap_{n,r}$, we have that for any $f \in C^{2}(\Xset)$ which satisfies Dirichlet boundary conditions (see footnote~\ref{footnote:dirichlet_eigenvalues}),
\begin{equation}
\label{eqn:seminorm_convergence}
\abs{f}_{H^1(\Xset)}^2 = \int_{\Xset} \bigl(\Delta_P f\bigr)(x) \cdot f(x) \,dx = \lim_{n \to \infty} \frac{1}{n^{2}r^{(d + 2)}}f^T \Lap_{n,r} f.
\end{equation}

\eqref{eqn:seminorm_convergence} establishes asymptotic equality (up to a scaling factor) between the penalties in~\eqref{eqn:smoothing_spline} and~\eqref{eqn:laplacian_smoothing}, when the distribution $P$ is uniform. But it also sheds light on an additional feature of Laplacian smoothing, relevant when $P$ is non-uniform: the penalty $f^T \Lap_{n,r} f$ automatically adapts to the design density $p$, in a way that the penalty $|f|_{H^1(\Xset)}^2$ does not. This automatic adaptivity to $P$ turns out to be a formal advantage of Laplacian smoothing relative to smoothing splines, as we will see in Section~\ref{sec:minimax_optimal_laplacian_smoothing}.

\paragraph{Minimax-optimality of smoothing splines.} 
One reason all of this is intriguing is that, as mentioned previously, the smoothing spline estimator $\wt{f}_{\SM}$---and the resultant test statistic $T_{\SM} := \|\wt{f}_{\SM}\|_n^2$--- have strong statistical properties. To be precise, they achieve the well-known minimax rates over Sobolev spaces.

Under~\ref{asmp:domain}, the minimax estimation rate over Sobolev balls is (see e.g. \citep{tsybakov2008_book})
\begin{equation}
\label{eqn:sobolev_space_estimation_minimax_rate}
\inf_{\wh{f}} \sup_{f_0 \in H^1(\Xset, M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp M^{2d/(2 + d)}n^{-2/(2 + d)}
\end{equation}
In the goodness-of-fit testing problem, we ask for a test function which can distinguish between the null and alternative hypotheses
\begin{equation}
\mathbf{H}_0: f_0 = 0, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{F} \setminus \{0\}
\end{equation} 
where $\mc{F}$ is some class of functions. The minimax critical radius is the smallest value of $\epsilon$ such that some level-${\alpha}$ test $\phi$ has power at least $1 - \alpha$ over all $\mc{F}_{\epsilon} := \mc{F} \cap \{f: \|f\|_{\Leb^2(\Xset)} \geq \epsilon\}$. Testing whether a regression function $f_0$ is equal to $0$ is an easier problem than estimating $f_0$, and so the minimax testing critical radius over $H^1(\Xset,M)$ is much smaller than the minimax estimation rate:
\begin{equation}
\label{eqn:sobolev_space_testing_critical_radius}
\epsilon^2\Bigl(H^1(\Xset,M)\Bigr):= \inf\Biggl\{\epsilon^2: \inf_{\phi} \biggl[\Ebb_0[\phi] +  \sup_{f_0 \in H_{\epsilon}^1(\Xset,M)} \Ebb_{f_0}[1 - \phi]\biggr] \leq 2\alpha \Biggr\} \asymp M^{2d/(4 + d)}n^{-4/(4 + d)}~~\textrm{for $1 \leq d < 4$;}
\end{equation}
see \cite{ingster09} for a proof of this statement, and \cite{ariascastro2018} for a more extended discussion of the minimax paradigm in nonparametric testing. When $d \geq 4$ we no longer have a continuous embedding of $H^1(\Xset)$ into $\Leb^4(\Xset)$; the functions in $H^1(\Xset)$ are very irregular, and the minimax rates in this regime are unknown.

>>>>

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

@@

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}

@@

\begin{document}
\title{Minimax optimal Laplacian smoothing}
\author{Alden Green}
\date{\today}
\maketitle

\textbf{Disclaimer:} This largely follows the same setup as in the draft I have written which combines Laplacian smoothing and Laplacian eigenmaps results. That draft --- document name \textit{graph\_regression2} --- should for the moment be considered obsolete.

@@

\textbf{(1)} Laplacian smoothing is a graph-based approach to nonparametric regression.

\begin{itemize}
	\item In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Rd$, and 
	\begin{equation}
	\label{eqn:random_design_regression}
	Y_i = f_0(X_i) + \varepsilon_i
	\end{equation}
	with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise. Our goal is to perform statistical inference on the unknown regression function $f_0: \Xset \to \Reals$, by which we mean either (a) \emph{estimating} $f_0$ by $\wh{f}$, an estimator constructed from the data $(X_1,Y_1),\ldots,(X_n,Y_n)$ or (b) simply \emph{testing} whether $f_0 = 0$, i.e whether there is any signal present. 
	\item In graph-based nonparametric regression, we perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense, and then constructing an estimate $\wh{f}$ which is smooth with respect to the graph $G_{n,r}$. The neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$. The $n \times n$ weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ encodes proximity between pairs of samples; for a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the entries $\mathbf{W}_{ij}$ are given by
	\begin{equation*}
	\label{eqn:neighborhood_graph}
	{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\Biggr).
	\end{equation*}
	Then the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$, and the graph Laplacian can be written as
	\begin{equation}
	\label{eqn:graph_Laplacian}
	\Lap_{n,r} = \bf{D} - \bf{W}
	\end{equation}
	\item The Laplacian smoothing estimator $\wh{f}_{\LS}$ \citep{smola2003} is a penalized least squares estimator, given by
	\begin{equation}
	\label{eqn:laplacian_smoothing}
	\wh{f}_{\LS} := \min_{f \in \Reals^n} \biggl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_{n,r} f \biggr\}
	\end{equation}
	where $\rho > 0$ acts a tuning parameter on the penalty $f^T \Lap_{{n,r}} f$.
	Assuming~\eqref{eqn:laplacian_smoothing} is a reasonable estimator of $f_0$, the squared empirical norm
	\begin{equation}
	T_{\LS} = \frac{1}{n}\Bigl\|\wt{f}_{\LS}\Bigr\|_2^2 \label{eqn:laplacian_smoothing_test}
	\end{equation}
	is in turn a reasonable statistic to assess whether or not $f_0 = 0$. 
\end{itemize}

\textbf{(2.a)} Laplacian smoothing has many advantages compared to other nonparametric regression methods. For instance:

\begin{itemize}
	\item It is fast, easy, and stable to compute. Letting ${\bf Y} = (Y_1,\ldots,Y_n) \in \Reals^n$, it is not hard to see that the minimizer $\wt{f}_{\LS} = (\rho \Lap_{n,r} + I)^{-1}{\bf Y}$ can be computed by solving a linear equation; in practice, one typically chooses $r$ to be small, with the result being that the matrix $\Lap_{n,r}$ is sparse. There exist known fast solvers of exactly this system.
	\item Is is generalizable to non-standard data---for instance, text or images, or really any data modality on which it is possible to define a kernel.
	\item It is easily adaptable to the semi-supervised learning setting. 
\end{itemize}

\textbf{(2.b)} For this reason a considerable body of work has emerged analyzing the consistency of the graph Laplacian $\Lap_{{n,r}}$ as $n \to \infty$ and $r(n) \to 0$. This is meant in various senses:
\begin{itemize}
	\item Pointwise consistency, meaning $\Lap_{n,r}f \to \Delta_Pf$ for an appropriate limiting operator $\Delta_P$.
	\item Spectral consistency, meaning the eigenvalues $\lambda_k(\Lap_{n,r}) \to \lambda_k(\Delta_P)$.
	\item Consistency of norms, meaning $f^T \Lap_{n,f} f \to \langle f,\Delta_Pf \rangle_{\Leb^2(P)}$. 
\end{itemize}
However, the statistical optimality of Laplacian smoothing is still not well understood.

\textbf{(2.c)} Our main contributions lie in this latter direction, and they can all be summarized as follows: Laplacian smoothing methods are minimax optimal over Sobolev spaces. In more detail, we will show that when $f_0$ belongs to the Sobolev ball $H^1(\Xset;M)$:
\begin{itemize}
	\item \textbf{Minimax optimal estimation.} With high probability, $\frac{1}{n}\norm{\wt{f}_{\LS} - f_0}_{2}^2 \lesssim n^{-2/(2 + d)}$.
	\item \textbf{Minimax optimal testing.}
	A level-$\alpha$ test constructed using $T_{\LS}$ has non-trivial power whenever $\norm{f_0}_{\Leb^2(\Xset)}^2 \gtrsim n^{-4/(4 + d)}$. 
	\item \textbf{Manifold adaptivity.}
	If $\mc{X} \subset \Rd$ is a submanifold of dimension $m < d$, both of the aforementioned rates hold with $d$ replaced by $m$.
\end{itemize}
Finally, we will also establish that the Laplacian smoothing estimator can be altered to take advantage of higher-order smoothness assumptions on $f_0$. This last conclusion will hold only under very limited circumstances.

\paragraph{Organization.}

List out what is to come.

\section{Minimax-optimal regression over Sobolev spaces}

Before we get to our main results, we briefly review minimax optimal estimation and testing rates over Sobolev classes. We also single out a notable method which achieves these rates: \emph{smoothing splines}. As we will see, Laplacian smoothing can be seen as a graph-based---i.e. discrete---counterpart to the ``continuous-time'' approach of smoothing splines.

\subsection{Sobolev spaces}
\label{subsec:sobolev_spaces}

We start by briefly introducing Sobolev spaces, emphasizing only the material relevant to our statistical context. Roughly speaking, Sobolev spaces contain functions $f \in \Leb^p(\Xset)$ with derivatives which themselves belong to $\Leb^p(\Xset)$. More formally, for given integers $s$ and $p > 0$, the Sobolev space $W^{s,p}(\Xset)$ consists of all functions $f \in \Leb^p(\Xset)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d) \in \mathbb{N}^d$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ exists and belongs to $\Leb^p(\Xset)$. For such functions, the $(s,p)$-Sobolev seminorm $\seminorm{f}_{W^{s,p}(\Xset)}$ and norm $\norm{f}_{W^{s,p}(\Xset)}$ are given by 
\begin{equation*}
\seminorm{f}_{W^{s,p}(\Xset)}^p = \sum_{\abs{\alpha} = s}\int_{\mathcal{X}} \Bigl|\bigl(D^{\alpha}f\bigr)(x)\Bigr|^p \,dx, ~~ \norm{f}_{W^{s,p}(\Xset)}^p = \sum_{k = 0}^{s} \seminorm{f}_{W^{k,p}(\Xset)}^p
\end{equation*}
and for a given $M > 0$, the $(s,p)$-Sobolev ball is $W^{s,p}(\Xset, M) = \set{f: \norm{f}_{W^{s,p}(\Xset)} \leq M}$. In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space, and we adopt the usual convention of writing $H^s(\Xset) = W^{s,2}(\Xset)$ and $H^s(\Xset,M) = W^{s,2}(\Xset,M)$; we will confine our attention to this case hereafter.

\textbf{(4)} In fact, an alternative way to motivate Laplacian smoothing nonparametric regression methods---as opposed to the advantages mentioned in \textbf{(2.a)}---comes from viewing them as discrete-time versions of smoothing splines. This is made more rigorous by the consistency results of \textbf{(2.b)}. Viewed in this light, an additional advantage of Laplacian smoothing emerges: it automatically adapts to the geometry of $\Xset$ and $p$.

\textbf{(5)} Smoothing splines are known to have strong minimax optimal properties. Reasoning by analogy, we might hope that Laplacian smoothing would share these properties.

\section{Graph Sobolev classes}

\textbf{(6)} The minimax optimality of smoothing splines rests on two crucial facts: one, the a priori assumption that $|f|_{H^1(\Xset)}$ is bounded; two, an upper bound on the metric entropy of $H^1(\Xset)$, characterized by the decay of its eigenvalues. \textcolor{red}{(TODO)}:(I know this isn't well-stated, and I need to clean it up.)  We now show that qualitatively similar statements hold with respect to

\quad \textbf{(6.1)} the graph Sobolev semi-norm $f_0^T \Lap_{n,r} f_0$ and,

\quad \textbf{(6.2)} the graph eigenvalues $\lambda_k(\Lap_{n,r})$.

\section{Minimax optimality of Laplacian smoothing}

\textbf{(7)} Using \textbf{(6.1)} and \textbf{(6.2)}, we establish upper bounds on the estimation and testing error of Laplacian smoothing, which confirm that Laplacian smoothing is asymptotically minimax rate-optimal over the Sobolev class $H^1(\Xset)$.

\textbf{(8)} While we motivated graph Laplacians by suggesting they might replicate the statistical properties of smoothing splines, we have in fact managed to show something stronger. In particular:

\quad \textbf{(8.1)} When $d \geq 2$, there is no Sobolev embedding of $H^1(\Xset) \subseteq C^s(\Xset)$ for any $s > 0$. Therefore the smoothing spline estimator is ill-posed.

\quad \textbf{(8.2)} Additionally, the metric entropy of $C^1(\Xset)$ is too large to obtain optimal rates for \textcolor{red}{Holder-smoothing splines} when $d \geq 2$. Thus, we see a gap emerge between the optimality properties of the discrete-time (Laplacian smoothing) and continuous-time (smoothing splines) estimators, in favor of the latter.

\quad \textbf{(8.3)} On the other hand, when $d \geq 4$, we no longer get the known optimal rates in the estimation problem. When $d = 4$, our rates are suboptimal by a factor of $\log n$. When $d > 5$ they are very suboptimal. We believe this is related to classical results regarding the entropy of low-smoothness Holder classes, i.e. the results used to justify \textbf{(8.2)}. The interesting point is that the parameter spaces $C^1(\Xset)$ and $H^1(G_{n,r})$ become ``too large'' for different values of $d$.

\quad \textbf{(8.4)} \textcolor{red}{(TODO)}: Translate these results into error in $\Leb^2(P)$ error, if possible.

\subsection{Manifold adaptivity}

\quad \textbf{(9)} We now show that when $\Xset$ is an $m$-dimensional submanifold of $\Reals^d$, Laplacian smoothing approaches automatically ``feel'' the intrinsic dimension $m$ of $\Xset$, which translate into improved upper bounds on estimating and testing error. This confirms the statement in \textbf{(4)} that Laplacian smoothing automatically adapts to the geometry of $\Xset$ and $p$.

\subsection{Higher-order smoothness classes}

\quad \textbf{(10)} Analogously to smoothing splines, Laplacian smoothing can also be adapted to take advantage of additional assumed regularity on $f_0$, i.e. $f_0 \in H^s(\mc{\Xset})$ for $s > 1$. In very limited circumstances, we can show that this adapted estimator achieves the sharper minimax rates over these classes. However, the general story--for all combinations of $s$ and $d$---remains beyond our reach.

\section{Simulations}

\quad \textbf{(11)} Empirically, we demonstrate that the risk for Laplacian smoothing is comparable to that of smoothing splines, in the full dimensional case. Although one cannot carry out a comparison in the case of an unknown manifold --- since the smoothing spline estimator must be changed in a non-obvious manner --- we do show that when the density $p$ is concentrated close to a manifold and $f_0$ is smooth where $p$ is peaked, we have improved performance for $\wh{f}_{\LS}$ compared to $\wh{f}_{\mathrm{SM}}$.

\section{Discussion}
